---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# safetymeta

<!-- badges: start -->
[![R-CMD-check](https://github.com/pharmacologie-caen/safetymeta/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/pharmacologie-caen/safetymeta/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The goal of safetymeta is to help manage adverse events aggregated datasets in order to perform
a safety meta-analysis.

## Installation

You can install the development version of safetymeta from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("pharmacologie-caen/safetymeta")
```

## Pre-requisites

Users are assumed to be familiar with

-    The principles of meta-analysis.

-    R (basic knowledge, a good start can be found [here](https://r4ds.hadley.nz/)).

-    The [PRISMA guidelines](https://www.prisma-statement.org/).

## What the package does not cover

-    Safety meta-analysis of individual patient level data.

## Overview

In this package, you will find steps to manage adverse events datasets,
avoiding common pitfalls in order to provide accurate, reproducible workflows
when conducting a safety meta-analysis.

# Usage

```{r}
library(safetymeta)
```

## Define priority among safety data sources

```{r sources}
source_order <- 
  c("src1", "src2")

e_data_source <-
  e_data |> 
  sort_sources(
    method = "source_list",
    source_name = "source",
    source_list_order = source_order
  )
```

## Group events into terms

```{r terms}
e_data_source |> 
  group_events(
    t_groups = t_groups,
    n_event = "n_event_arm1",
    n_arm = "n_arm1",
    event_name = "event_name",
    study_id = "study_id"
  )
```


## Vignettes

Look at the "basic_workflow" article for more details.
