---
title: "basic_workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{basic_workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, warning=FALSE}
library(safetymeta)
library(dplyr)
```

# Purpose

To help manage adverse events aggregated datasets in order to perform
a safety meta-analysis.

## Pre-requisites

Users are assumed to be familiar with

-    The principles of meta-analysis.

-    R (basic knowledge, a good start can be found [here](https://r4ds.hadley.nz/)).

-    The [PRISMA guidelines](https://www.prisma-statement.org/).

## What the package does not cover

-    Safety meta-analysis of individual patient level data.

## Overview

In this package, you will find steps to manage adverse events datasets,
avoiding common pitfalls in order to provide accurate, reproducible workflows
when conducting a safety meta-analysis.

# Usage

## Gather safety data from different sources

According to PRISMA Guidelines, events identification required to screen
all potential sources of information. This includes direct contact to the authors,
published data from original manuscript, its supplementary data, potential follow-up studies,
and clinical trial registration websites, among others.

Not all sources will provide data for each study, and not all sources will
provide the **same** data for each study. A choice is required in the latter case,
in order to retain the most relevant source.

When we have to choose a preferred source, several considerations may come into mind:

-   How reliable is the source? Which source should be trusted the most?

If there was a straightforward answer to this question, there would be no other 
discussion, and sources would be prioritized according to truthfulness.
It happens to be a not so trivial question, as no guidelines establish a
hierarchical order among sources. Arguably, not all sources would be as valuable
for all research questions as well.

Other, more simple features, can also help:

-   Which source reports on the highest number of patients?

-   Which source has the highest number of events?

-   Which source has the longest follow-up?

The process of sorting sources is topic-dependent and likely meta-analysis-dependent,
which means sorting would not be identical for 2 different meta-analyses.

More importantly, authors should be aware that such sorting is required before
processing data. A dedicated time should be spend to determine the most
appropriate order.

Here is an example, with two different data sources

```{r data_sources}

e_data_src1

e_data_src2
```
As we can see, study with id "NCT2" is present in both sources. The same
event "eczema" is reported in both, but the number of patients presenting
this event (column `n_event_arm1`) is different: 10 patients in src1, vs 9 in src2

Also, sometime, src2 is the only source to provide data, such as for study "NCT3".

With source sorting, data are regarded step by step:

-    Does no 1 source has data? If yes, use this source for this event

-    If no 2 source doesn't provide data for this event, does no 2 source do? If yes, use no 2.

-    Etc.

This also implies that if you study different events, you will not always
use the same source.

### Sort sources by yourself

If it has been decided which source should be prioritize, you can give this data
to the `sort_sources()` function.

```{r source_order}

e_data <- 
  bind_rows(
    e_data_src1,
    e_data_src2
    )

source_order <- 
  c("src1", "src2")

```

The `sort_sources()` function will take a `source_list_order` argument, which tells us
the priority among sources. The first item is the source with the highest priority,
the second element is the next highest priority, and so on.

```{r sort_sources}
e_data_source <-
  e_data |> 
  sort_sources(
    method = "source_list",
    source_name = "source",
    source_list_order = source_order
  )
```

All the function does is adding an integer column "src" to the initial dataset.

This will be helpful for the next step, gathering safety data.

## Gather events into terms

Once sources have been sorted, events can be gathered.

A same medical condition can sometime be defined by several names.

Here, we use the word `term` to refer to a medical condition that can have several
names. Those different names are called `events`.

-    An "event" dataset contains one row per reaction, with counts of events per arm,
n per arm, and a study_id column (see example).

-    A "term" is a combination of one or more "events" (can be from an SMQ,
 an HLT, any combination of PT in MedDRA, ...).
 
 Lets take the medical condition "myocardial infarction". You could say the patient
 had myocardial infarction if a study reports on myocardial infarction, but also 
 on Sustained ST-elevation myocardial infarction, or acute coronary syndrome.
 All these *events* can define the *term* "myocardial infarction".
 In different studies (or in different centers in the same study), the term maybe
 reported with one of these events interchangeably. Or some patients could be
 reported as one event, and others as another.
 
 To capture all these events, setting-up a dictionary of terms is required.
 
```{r t_groups_mi}
t_groups <- 
  list(myocardial_infarction = 
         c("myocardial infarction",
           "sustained st-elevation myocardial infarction",
           "acute coronary syndrome")
  )

t_groups
```
We now have a named list, containing one element. This element is named after the
**term** myocardial infarction, and gathers all possible **events**.
 
We can use the `group_events()` function to reorganize our dataset
from events to terms.

```{r group_events_basic}

e_data_source |> 
  group_events(
    t_groups = t_groups,
    n_event = "n_event_arm1",
    n_arm = "n_arm1",
    event_name = "event_name",
    study_id = "study_id"
  )

```

Two interesting behaviors here:

### The e_data dataset has been turned into a **term**-level dataset.

Counts in n_event_arm1 have changed: instead of two lines for study NCT1, with
1 and 2 events in myocardial infarction and acute coronary syndrome, we now have
only one line, with 2 events.

This is the default behavior of the function, with method named "conservative".

The idea is that since events are reported as grouped counts, it is now always
possible to say if patient with a first event (e.g. myocardial infarction) did 
not also presented a second event (e.g. acute coronary syndrome).

In some special cases, you might want to sum up those counts. That's particularly
true if when working with rare events, or if you join events that should not
(in principle) overlap one with each other. For all those cases, the "integrative"
method will perform a sum.

```{r group_events_integrative}

e_data_source |> 
  group_events(
    t_groups = t_groups,
    n_event = "n_event_arm1",
    n_arm = "n_arm1",
    event_name = "event_name",
    study_id = "study_id",
    method = "integrative"
  )

```
For NCT1, there are now 3 events in arm 1.

### Data from most relevant source has been retained

Lets capture another term

```{r t_groups_skin}
t_groups <- 
  list(skin = 
         c("eczema",
           "pruritus")
  )

e_data_source |> 
  group_events(
    t_groups = t_groups,
    n_event = "n_event_arm1",
    n_arm = "n_arm1",
    event_name = "event_name",
    study_id = "study_id"
  )
```

In study NCT2, eczema was reported from 2 different sources (src1, and src2).

Since those sources were prioritized with `sort_sources()`, `group_events()` will only
keep data from src1.

Of important note, source filtering in `group_events()` is at term level.

```{r ge_ss_term_level}
# load data t_groups, with the 2 previous groups
data(t_groups)

 e_s <-
    data.frame(
      study_id = c("NCT1", "NCT1"),
      event_name = c("myocardial infarction",
                     "eczema"),
      src = c(1, 2),
      n_event_arm1 = c(2, 15),
      n_arm1 = c(20, 20)
    )

e_s |>
    group_events(
      t_groups = t_groups,
      n_event = "n_event_arm1",
      n_arm = "n_arm1",
      event_name = "event_name",
      study_id = "study_id"
    )


```

In this new example, there is only one study, two events reported from two sources.

Since these events are identified each in one term from t_groups, and each from 
the source with the highest priority (source 1 for term "myocardial_infarction",
source 2 for term "skin"), there are both retained.
