---
title: "collect_data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{collect_data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(safetymeta)
library(meta)
```

Once safety meta-analysis has been performed, it is natural to collect numerical
results on the model(s).

Although not specific of safety meta-analysis, the package provides the 
`table_mod_summary()` function to table those summary results.

This is especially useful when there are multiple models, e.g, when studying
more than one event, or performing subgroup analyses.

```{r tms_onemod}
data <-
    data.frame(
      event_a = 4:1,
      event_b = 1:4,
      n = c(5, 7, 8, 9),
      byvar = c("a", "a", "b", "b")
    )

 m_a <- metaprop(
    event = event_a,
    n = n,
    byvar = byvar,
    data = data
  )
 
 

table_mod_summary(m_a)


```
If studying two events, use a named list of models

```{r tms_twoevents}

m_b <- metaprop(
    event = event_b,
    n = n,
    byvar = byvar,
    data = data
  )


table_mod_summary(
     list(mod_a = m_a,
          mod_b = m_b)
   )
```
Notice that the `names` of the model `list` are collected into the `mod_name`
column. Providing enough data in these names makes it easy to track the results
in subsequent data management steps.

If studying subgroups, use argument `main_or_sbgp`.

```{r tms_sbgp}

m_a |> 
  table_mod_summary(
    main_or_sbgp = "sbgp"
  )
```
> For `meta::metaprop` models, `table_mod_summary()` only works if `sm="PLOGIT"`

## Multiple subgroups

You can combine `metabin_wrap()` to `table_mod_summary()`

```{r sbgp}

tnsc_data |>
  dplyr::filter(term_name == "myocardial_infarction") |>
  metabin_wrap(
    event_colnames = c("event_e", "event_c"),
    n_colnames = c("n_e", "n_c"),
    by_var = c("byvar1", "byvar2"),
    stud_lab = "study_id"
  ) |> 
   table_mod_summary(
    main_or_sbgp = "sbgp"
  )

```

